 

### 2020-09-29 vertical-thermal-review.Rmd
Update vertical-thermal-review.Rmd

**Air temperature, $T_{air}$, is sometimes significantly buffered by forest canopies (Fig. **2x**).**
Studies comparing $T_{air}$ under forest canopies with nearby clearings have found lower maximum temperatures under forest canopies across Europe [@zellweger_seasonal_2019] and in the northwestern United States [@davis_microclimatic_2019]. 
Similarly, maximum air temperatures were higher above than below tropical forest canopies during wet seasons in Panama [@rey-sanchez_spatial_2016] and coastal Brazil [@fauset_differences_2018, Tymen et al. 2017] 
"max temps were higher and humidity lower in gaps vs understory regions (both at 1.5 m above the ground), see Fig. 4 - again illustrating buffering effect of the canopy"
However, similar maximum temperatures have been observed during the dry season in Panama [@rey-sanchez_spatial_2016] and in a temperate deciduous forest in the eastern United States [@mcgregor_tree_nodate].
Minimum $T_{air}$ is also buffered by forest canopies under some conditions. 
Higher minimum temperatures under forest canopies relative to nearby clearings have been observed across Europe [@zellweger_seasonal_2019] and in the northwestern United States [@davis_microclimatic_2019].
However, similar minimum temperatures under forest canopies have been observed in tropical forests in Panama [@rey-sanchez_spatial_2016] and coastal Brazil [@fauset_differences_2018], as well as in a temperate deciduous forest in the eastern United States [@mcgregor_tree_nodate].
*(Add something about risk of freeze.)*

**Humidity also varies across the forest vertical profile, being generally higher in the understory (Fig. **2x**).**
- DETAILS - REFS
- @mcgregor_tree_nodate
RH, in combination with temperature, determines vapor pressure deficit, $VPD$
- $VPD$ is lower in understory of Atlantic forest during wet season [@fauset_differences_2018, Tymen et al. 2017]
This means that canopy leaves tend to be exposed to higher *evaporative demand*.

@@

*Leaf area / characteristic dimension*:
- Recent paper on influence on leaf size/shape in PCE – certain aspects of leaf shape were not as relevant as expected. 
- Max stomatal conductance - increases with canopy height (Kenzo et al. 2015).
Max transpiration rate - increases with canopy height (Kenzo et al. 2015).

>>>>

**Air temperature, $T_{air}$, is sometimes significantly buffered by forest canopies (Fig. **2x**).**
Studies comparing $T_{air}$ under forest canopies with nearby clearings have found lower maximum temperatures under forest canopies across Europe [@zellweger_seasonal_2019, @vonarxSpatiotemporalEffectsForest2012] and in the northwestern United States [@davis_microclimatic_2019]. 
Similarly, maximum air temperatures were higher above than below tropical forest canopies during wet seasons in Panama [@rey-sanchez_spatial_2016] and coastal Brazil [@fauset_differences_2018, Tymen et al. 2017] 
"max temps were higher and humidity lower in gaps vs understory regions (both at 1.5 m above the ground), see Fig. 4 - again illustrating buffering effect of the canopy"
However, similar maximum temperatures have been observed during the dry season in Panama [@rey-sanchez_spatial_2016] and in a temperate deciduous forest in the eastern United States [@mcgregor_tree_nodate].
Minimum $T_{air}$ is also buffered by forest canopies under some conditions. 
Higher minimum temperatures under forest canopies relative to nearby clearings have been observed across Europe [@zellweger_seasonal_2019] and in the northwestern United States [@davis_microclimatic_2019].
However, similar minimum temperatures under forest canopies have been observed in tropical forests in Panama [@rey-sanchez_spatial_2016] and coastal Brazil [@fauset_differences_2018], as well as in a temperate deciduous forest in the eastern United States [@mcgregor_tree_nodate, @nakamuraForestsTheirCanopies2017].
*(Add something about risk of freeze.)*--> [@nakamuraForestsTheirCanopies2017]

**Humidity also varies across the forest vertical profile, being generally higher in the understory (Fig. **2x**).**
- DETAILS - REFS
- @mcgregor_tree_nodate

Forest canopies moderate relative humidity in below-canopy microclimates along with temperature. In three European forest types (Broadleaved, Non-Pine Conifer, Pine) canopy buffering was observed to decrease daily maximum $T_{air}$ and increase daily minimum RH in below-canopy microclimate [@vonarxSpatiotemporalEffectsForest2012]

RH, in combination with temperature, determines vapor pressure deficit, $VPD$
- $VPD$ is lower in understory of Atlantic forest during wet season [@fauset_differences_2018, Tymen et al. 2017]
This means that canopy leaves tend to be exposed to higher *evaporative demand*, higher stomtal contrains and lower osmotic potential, making canopy leaves more efficient in regulating water use [@niinemetsPhotosyntheticAcclimationSimultaneous2004]

@@

*Leaf area / characteristic dimension*:
- Recent paper on influence on leaf size/shape in PCE – certain aspects of leaf shape were not as relevant as expected. 
- Max stomatal conductance - increases with canopy height (Kenzo et al. 2015)
Max transpiration rate - increases with canopy height (Kenzo et al. 2015).
